Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-06-15 WRITEUP/ms.tex
minor comments added; i would substitute figure 3 and 4 with 5 (it is packed version)

\begin{figure}[htbp]
\centering
\includegraphics[trim=0 0 0 0, clip, width=0.5\textwidth]{\figpath/hist_test_16.pdf}
\caption{Distribution of molecular gas density of \flower during the accreting phase shown in \Fig{SFH}{\em
    (b)}.
\label{fig:h2density}}
\end{figure}

% This is the messy stage
\begin{figure*}[htbp]
 \centering
  \includegraphics[scale=0.6]{\figpath/{dual_16_ncut_0.53}.pdf}
  \\ [-2.9em]
  \includegraphics[scale=0.6]{\figpath/{dual_16_ncut_6.81}.pdf}
  \\ [-2.9em]
  \includegraphics[scale=0.6]{\figpath/{dual_16_ncut_18.96}.pdf}
\caption{
Examples of MCCs (white contours) identified by the clump finder in
\flower during its accreting phase (\Fig{SFH}a), where \flower is dispersion-dominated.
Colorbar on the right shows the mean
H$_2$ number density, weighted by gas mass. Different rows show the results obtained by applying different H$_2$ number density cuts $(n_{\rm cut})$ as shown by the label. Left and right panels show the galaxy from different viewing angles.
\label{fig:MCC}}
\end{figure*}

% Add a figure that shows \flower in "disky" stage. We are interested in exploring whether the MCC properties are dependent on the disk properties.
% In other words, if the nature of the MCCs we are identifying are large chucks of arms/disk in \flower, then its dynamics that we are deriving will be dominated by
% large-scale shear.
\begin{figure*}[htbp]
 \centering
  \includegraphics[scale=0.6]{\figpath/{dual_28_ncut_0.53}.pdf}
  \\ [-2.7em]
  \includegraphics[scale=0.6]{\figpath/{dual_28_ncut_6.81}.pdf}
  \\ [-2.7em]
  \includegraphics[scale=0.6]{\figpath/{dual_28_ncut_18.96}.pdf}
\caption{
Same as \Fig{MCC}, except we show the quiescent phase of \flower, which is supported by rotation.
\label{fig:MCC28}}
\end{figure*}

@@

\subsection{Distributions of MCC Basic Properties and Outlier} \label{sec:dist}
\begin{figure*}[htbp]
\centering
\includegraphics[trim=0 0 0 0, clip, width=\textwidth]{\figpath/minmaxNcut_basicDistributions.pdf}
\caption{Distributions of mass (left), size (middle), and gas mass fraction (right) of MCCs identified using the lowest 
$n_{\rm cut}$\eq0.32\,\cc (top panels) and $n_{\rm ncut}$\eq18.96\,\cc (bottom panels) over all the considered evolutionary stages of \flower traced in the simulation. Note that the scales shown on the $y$-axes are different between the top and bottom panels, as fewer MCCs are identified at higher $n_{\rm cut}$.
\label{fig:dist}}
\end{figure*}

@@

% address the biggest MCC = molecular disk of \flower
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{\figpath/cloud-mass_stellar-to-gas-mass_ss16.png}
\caption{Cloud mass and stellar-to-gas mass ratio of MCCs identified in the accreting phase of \flower with different $n_{\rm cut}$.
The most massive MCC identified at low $n_{\rm cut}$ shows 
a high stellar-to-gas mass ratio ($\gtrsim$\,1) \AP{do we really need both $f_{\rm gas}$ and $M_\star/M_{\rm gas}$?} 
and corresponds to the molecular disk of \flower (see also \Fig{vv}). Thus,
it includes a large number of stellar mass that is already assembled in \flower. Such structure is excluded 
in the discussion of MCC dynamics in the remainder of this paper.
\label{fig:stellarRatio16}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[trim=0 0 0 0, clip, width=0.5\textwidth]{\figpath/ss16-sigma-kms-bulk_sigma-kms-NT} \\
\includegraphics[trim=0 0 0 0, clip, width=0.5\textwidth]{\figpath/ss28-sigma-kms-bulk_sigma-kms-NT}
\caption{Gas velocity dispersions resulting from bulk versus turbulent motions 
for MCCs identified across all $n_{\rm cut}$ (same color-coding as \Fig{stellarRatio16}) in the accreting (top) and
quiescent phase (bottom) of \flower.
The accreting phase (\Fig{MCC}) of \flower displays a more disturbed morphology compared to the
quiescent phase, which displays a disk-like morphology (\Fig{MCC28}).
Velocity dispersion from turbulent motions dominates over bulk motions for most MCCs in the accreting phase. 
On the other hand, $\sigma_{\rm bulk}$ is at least comparable to $\sigma_{\rm NT}$ for almost half the MCCs
in the quiescent disk-like phase.
\label{fig:vv}}
\vspace{0.5em}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[trim=0 0 0 0, clip, width=0.92\textwidth]{\figpath/{ss16_toomre_combined_3by1_0_2.0}.pdf}
\includegraphics[trim=0 0 0 65, clip, width=0.95\textwidth]{\figpath/{ss16_toomre_combined_3by1_0_0.8}.pdf}
\caption{
Toomre $Q$ maps derived from the
central $r$\eq2\,kpc (top row) and $r$\eq0.8\,kpc (bottom row) of \flower.
Gas-only $Q_{\rm gas}$ is shown in the left panels and stellar-only $Q_\star$ is shown in the middle panels.
Maps of the effective two-component Toomre $Q_{\rm eff}$ parameter are shown in the right panels.
All maps are projected onto the $xy$-plane.
Positions of MCCs identified with $n_{\rm cut}$\eq6.81\,cm$^{-3}$ are overplotted as star symbols.
A smoothing length of 30\,pc has been applied to the maps.
A divergent colormap is used for the Toomre $Q$ maps to facilitate
identification of regions above and below $\log{Q}$\eq0.
Some MCCs lie in regions of $\log{Q_{\rm eff}}\gtrsim0$, where regions of $\log{Q_{\rm eff}}\lesssim0$
are likely gravitationally unstable.
Close resemblance of the $Q_\star$ and $Q_{\rm eff}$ maps in the {\em central region} of \flower
indicates that the stellar component plays an important role in
governing the stability of the MCCs against $m=0$  perturbations,
highlighting the importance of accounting for stellar contribution when examining the stability of molecular gas in relatively evolved and enriched systems at high redshifts. \AP{maybe it is better to put a single set of maps, with FOV from -1.5 to 1.5 kpc}
\label{fig:Qeff}}
\end{figure*}

@@

Properties of the star-forming ISM of galaxies near the Epoch of Reionization are now within reach with ALMA. While it is possible to obtain sensitive and high fidelity imaging that reveals their gas and dust morphology on sub-kpc scales and even smaller, such \obs remain time-consuming; logistically, data from multiple cycles pushing to increasing resolution and sensitivity are needed. As such, observational studies of MCCs on such scales are still missing in the literature.
%
In this work, we aim to understand the origin and dynamical properties of MCCs in prototypical galaxies at the EoR in numerical simulations to provide a framework within which upcoming \obs can be compared against to aid in the interpretation.

We study the dynamics of MCCs in a \z$\sim$\,6 prototypical galaxy, \flower, and their temporal evolution using the state-of-the-art cosmological zoom-in simulations (\ncode{Serra}), which include a chemical network to determine the formation of molecular hydrogen, heating and cooling of the ISM by UV radiation and metal lines, and detailed stellar feedback.
%
We use a three-dimensional clump-finding algorithm to identify MCCs. We decompose the molecular structures into non-overlapping objects by identifying a set of H$_2$ density contours with varying densities ($n_{\rm cut}$) at multiple evolutionary stages. Most notably, the accreting phase, starburst phase, and quiescent phase. We extract properties such as mass, size, Mach number, velocity dispersion, gas surface density, and virial parameter ($M_{\rm gas}, R, \mathcal{M}, \sigma_{\rm gas}, \Sigma_{\rm gas}, \alpha_{\rm vir}$) for each MCC and perform a Toomre-$Q$ stability analysis on \flower.

@@

The variation in $\sigma_{\rm gas}^2/R$ with $\Sigma$ seen in this work is in line with previous studies, suggesting that ``Larson's third relation'' is an artifact arising from observational bias artifact of the limited range of  column densities a specific molecular line tracer is sensitive to \citep[][]{Ballesteros02a, Ballesteros11a}. This would imply column density is not a fundamental properties of molecular clouds and not all molecular clouds are virialized.

@@

% Toomre:
Close resemblance of $Q_{\rm eff}$ and $Q_{\rm star}$ maps indicates that contribution from the stellar component plays an important role in governing the stability of the MCCs against axisymmetric perturbations, especially in the central part of \flower. Similarly, stabilizing effect due to the thickness of its disk is also non-negligible. This illustrates the importance of accounting for stellar contribution and disk thickness when examining the stability of molecular gas structures, especially in relatively evolved and enriched systems at high redshift that are preferentially being observed now.

% where the virial parameter is used as a surrogate for comparing the potential to kinetic energies of molecule gas structures. 
We perform virial analysis, as motivated by \obs, to assess the stability of MCCs. On average, the virial parameter of MCCs in the starburst phase of \flower is lower than in the accreting phase, as expected for \SF. Similarly, the virial parameter of MCCs identified at the highest density thresholds are lower than those identified at lower 
density thresholds. 
% Such MCCs are thus susceptible to collapse. 
% The high virial parameter found for some MCCs are found in regions of $Q_{\rm eff}>1$.
% On the other hand, 
Such MCCs identified in (the denser) regions with a lower Toomre $Q_{\rm eff}$ parameter are unstable against collapse. Star formation is expected to take place within its star-forming {\it clumps} and {\it cores} on smaller scales, where energy quickly dissipates. This is consistent with the notion that collapsing structures result from gravitational instability occurring within globally stable structures, which are supported by turbulence and rotation on large scale.
%%
%Low-$Q$ sets the regions that fragments 
% and within those regions where alphavir is low enough is where structures to further collapse to form stars.
%
%The MCCs of \flower are more massive and turbulent compared to the Milky Way, with 
% higher gas mass surface densities and velocity dispersions. 
This also implies that \obs on spatially resolution better than $\simeq$40\,pc are needed to examine the truly star-forming structures (cores), and thus, \SF in the first galaxies.
%
%% end
%Determining the multi-phase ISM properties of early galaxies
%is a critical piece to understanding the evolution and
%assembly history of galaxies, since they set the pace
%for chemical reactions and excitation rates for the coolants in the ISM (and subsequent star formation).
%While observations leveraging the combination of spatial-spectral imaging of
%multi-band continuum and spectral line emission are crucial,
%% we show in this work that prototypical galaxies at the EoR with properties similar to \flower
%% are expected to have molecular gas structures on sub-kpc scales. 
% Higher resolution imaging of molecular gas 
Such resolution is within reach with ALMA. % 0.01" at the high freq bands
The Next Generation VLA (ngVLA) in the 2030s will enable deeper imaging of the cold gas down to $\sim$\,200\,pc via \aco line emission. % 0.005" highest freq; ~0.05 for 16 GHz; CO10 for z~6
On the other hand, cosmological zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy statistics and are subject to the sub-grid models adopted, serve as a useful tool for examining and making predictions on the morphology and dynamics of the molecular ISM of the first galaxies.
%These predictions can in turn be
%tested with high resolution imaging of the gas content in the first galaxies with telescopes such as ALMA and the ngVLA, thereby
%enabling us to test our findings and refine future simulations to shed light on the physical processes behind
%\SF since the cosmic dark ages.

>>>>

\begin{figure}[htbp]
\centering
\includegraphics[trim=0 0 0 0, clip, width=0.5\textwidth]{\figpath/hist_test_16.pdf}
\caption{Distribution of molecular gas density of \flower during the accreting phase shown in \Fig{SFH}{\em
    (b)} \AP{what is the normalization/axis here? I would plot the distribution such that the integral on the function is 1}.
\label{fig:h2density}}
\end{figure}

% This is the messy stage
\begin{figure*}[htbp]
 \centering
  \includegraphics[scale=0.6]{\figpath/{dual_16_ncut_0.53}.pdf}
  \\ [-2.9em]
  \includegraphics[scale=0.6]{\figpath/{dual_16_ncut_6.81}.pdf}
  \\ [-2.9em]
  \includegraphics[scale=0.6]{\figpath/{dual_16_ncut_18.96}.pdf}
\caption{
Examples of MCCs (white contours) identified by the clump finder in
\flower during its accreting phase (\Fig{SFH}a), where \flower is dispersion-dominated.
Colorbar on the right shows the mean
H$_2$ number density, weighted by gas mass. Different rows show the results obtained by applying different H$_2$ number density cuts $(n_{\rm cut})$ as shown by the label. Left and right panels show the galaxy from different viewing angles.
\label{fig:MCC}\AP{see \Fig{MMCidentification} for an alternative version}}
\end{figure*}

% Add a figure that shows \flower in "disky" stage. We are interested in exploring whether the MCC properties are dependent on the disk properties.
% In other words, if the nature of the MCCs we are identifying are large chucks of arms/disk in \flower, then its dynamics that we are deriving will be dominated by
% large-scale shear.
\begin{figure*}[htbp]
 \centering
  \includegraphics[scale=0.6]{\figpath/{dual_28_ncut_0.53}.pdf}
  \\ [-2.7em]
  \includegraphics[scale=0.6]{\figpath/{dual_28_ncut_6.81}.pdf}
  \\ [-2.7em]
  \includegraphics[scale=0.6]{\figpath/{dual_28_ncut_18.96}.pdf}
\caption{
Same as \Fig{MCC} for the quiescent phase of \flower, when the galaxy is supported by rotation.
\label{fig:MCC28} \AP{see \Fig{MMCidentification} for an alternative version}}
\end{figure*}

\begin{figure*}[htbp]
 \centering
  \includegraphics[width=0.49\textwidth]{\figpath/{dual_16_ncut_0.53}.pdf}
  \includegraphics[width=0.49\textwidth]{\figpath/{dual_28_ncut_0.53}.pdf}

  \includegraphics[width=0.49\textwidth]{\figpath/{dual_16_ncut_6.81}.pdf}
  \includegraphics[width=0.49\textwidth]{\figpath/{dual_28_ncut_6.81}.pdf}

  \includegraphics[width=0.49\textwidth]{\figpath/{dual_16_ncut_18.96}.pdf}
  \includegraphics[width=0.49\textwidth]{\figpath/{dual_28_ncut_18.96}.pdf}
\caption{\AP{proposed alternative to substitute \Fig{MCC} and \Fig{MCC28}; if adopted, label with the phase should be included in the first row}
\label{fig:MMCidentification}
}
\end{figure*}

@@

\subsection{Distributions of MCC Basic Properties and Outlier} \label{sec:dist}
\begin{figure*}[htbp]
\centering
\includegraphics[trim=0 0 0 0, clip, width=\textwidth]{\figpath/minmaxNcut_basicDistributions.pdf}
\caption{Distributions of mass (left), size (middle), and gas mass fraction (right) of MCCs identified using the lowest 
$n_{\rm cut}$\eq0.32\,\cc (top panels) and $n_{\rm ncut}$\eq18.96\,\cc (bottom panels) over all the considered evolutionary stages of \flower traced in the simulation. \AP{I would normalize the distributions such that their sum is 100\% (bin width already accounted for), so that the comparison is more direct; on a side, could we use a different shade of blue? (see also \Fig{h2density})} Note that the scales shown on the $y$-axes are different between the top and bottom panels, as fewer MCCs are identified at higher $n_{\rm cut}$.
\label{fig:dist}}
\end{figure*}

@@

% address the biggest MCC = molecular disk of \flower
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{\figpath/cloud-mass_stellar-to-gas-mass_ss16.png}
\caption{Cloud mass and stellar-to-gas mass ratio of MCCs identified in the accreting phase of \flower with different $n_{\rm cut}$. The most massive MCC identified at low $n_{\rm cut}$ shows a high stellar-to-gas mass ratio ($\gtrsim$\,1) \AP{do we really need both $f_{\rm gas}$ and $M_\star/M_{\rm gas}$?} and corresponds to the molecular disk of \flower (see also \Fig{vv}).
%
Thus, it encloses a large of stellar mass that is already assembled in \flower. The disk structure is excluded in the discussion of MCC dynamics in the remainder of this paper. \AP{minor mod: legend with horizontal alignment, text with arrows pointing to the disk and proper MMC clouds of points}
\label{fig:stellarRatio16}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[trim=0 0 0 0, clip, width=0.5\textwidth]{\figpath/ss16-sigma-kms-bulk_sigma-kms-NT} \\
\includegraphics[trim=0 0 0 0, clip, width=0.5\textwidth]{\figpath/ss28-sigma-kms-bulk_sigma-kms-NT}
\caption{Gas velocity dispersions resulting from bulk versus turbulent motions for MCCs identified across all $n_{\rm cut}$ (same color-coding as \Fig{stellarRatio16}) in the accreting (top) and quiescent phase (bottom) of \flower.
%
The accreting phase (\Fig{MCC}) of \flower displays a more disturbed morphology compared to the quiescent phase, which displays a disk-like morphology (\Fig{MCC28}).
%
Velocity dispersion from turbulent motions dominates over bulk motions for most MCCs in the accreting phase. On the other hand, $\sigma_{\rm bulk}$ is at least comparable to $\sigma_{\rm NT}$ for almost half the MCCs in the quiescent disk-like phase.
\label{fig:vv}}
\vspace{0.5em}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[trim=0 0 0 0, clip, width=0.92\textwidth]{\figpath/{ss16_toomre_combined_3by1_0_2.0}.pdf}
\includegraphics[trim=0 0 0 65, clip, width=0.95\textwidth]{\figpath/{ss16_toomre_combined_3by1_0_0.8}.pdf}
\caption{
Toomre $Q$ maps derived from the central $r$\eq2\,kpc (top row) and $r$\eq0.8\,kpc (bottom row) of \flower. The contribution from the gas ($Q_{\rm gas}$) and the stellar ($Q_\star$) components is shown separately in the left and middle panels, respectively. Maps of the effective two-component Toomre $Q_{\rm eff}$ parameter are shown in the right panels.
%
All maps are projected onto the plane of the disk, and a gaussian smoothing of width 30\,pc has been applied to the maps.
%
Positions of MCCs identified with $n_{\rm cut}$\eq6.81\,cm$^{-3}$ are overplotted as star symbols.
%A divergent colormap is used for the Toomre $Q$ maps to facilitate identification of regions above and below $\log{Q}$\eq0.
Some MCCs lie in regions of $\log{Q_{\rm eff}}\gtrsim0$, where regions of $\log{Q_{\rm eff}}\lesssim0$ are likely gravitationally unstable.
%
Close resemblance of the $Q_\star$ and $Q_{\rm eff}$ maps in the {\em central region} of \flower indicates that the stellar component plays an important role in governing the stability of the MCCs against $m=0$  perturbations, highlighting the importance of accounting for stellar contribution when examining the stability of molecular gas in relatively evolved and enriched systems at high redshifts. \AP{maybe it is better to put a single set of maps, with FOV from -1.5 to 1.5 kpc}
\label{fig:Qeff}}
\end{figure*}

@@

Properties of the star-forming ISM of galaxies near the Epoch of Reionization are now within reach with ALMA. While it is possible to obtain sensitive and high fidelity imaging that reveals their gas and dust morphology on sub-kpc scales and even smaller, such \obs remain challenging; logistically, data from multiple cycles pushing to increasing resolution and sensitivity are needed. As such, observational studies of MCCs on such scales are still missing in the literature.
%
In this work, we aim to understand the origin and dynamical properties of MCCs in prototypical galaxies at the EoR in numerical simulations to provide a framework within which upcoming \obs can be compared against to aid in the interpretation.

We study the dynamics of MCCs and their temporal evolution in \flower, a \z$\sim$\,6 prototypical galaxy obtained from the state-of-the-art cosmological zoom-in simulations \ncode{Serra}, which include a chemical network to determine the formation of molecular hydrogen, heating and cooling of the ISM by UV radiation and metal lines, and detailed stellar feedback.
%
We use a three-dimensional clump-finding algorithm to identify MCCs. We decompose the molecular structures into non-overlapping objects by using a set of H$_2$ density contours ($n_{\rm cut}$) at multiple evolutionary stages, in particular focusing on the accreting phase, starburst phase, and quiescent phase. We extract properties such as mass, size, Mach number, velocity dispersion, gas surface density, and virial parameter ($M_{\rm gas}, R, \mathcal{M}, \sigma_{\rm gas}, \Sigma_{\rm gas}, \alpha_{\rm vir}$) for each MCC and perform a Toomre-$Q$ stability analysis on \flower.

@@

\AP{the following should be removed if the corresponding point in the main text is commented out (as it has been now)}
The variation in $\sigma_{\rm gas}^2/R$ with $\Sigma$ seen in this work is in line with previous studies, suggesting that ``Larson's third relation'' is an artifact arising from observational bias artifact of the limited range of  column densities a specific molecular line tracer is sensitive to \citep[][]{Ballesteros02a, Ballesteros11a}. This would imply column density is not a fundamental properties of molecular clouds and not all molecular clouds are virialized.

@@

% where the virial parameter is used as a surrogate for comparing the potential to kinetic energies of molecule gas structures. 
We perform virial analysis, as motivated by \obs, to assess the stability of MCCs. On average, the virial parameter of MCCs in the starburst phase of \flower is lower than in the accreting phase, as expected for \SF. Similarly, the virial parameter of MCCs identified at the highest density thresholds are lower than those identified at lower density thresholds. 
% Such MCCs are thus susceptible to collapse. 
% The high virial parameter found for some MCCs are found in regions of $Q_{\rm eff}>1$.
% On the other hand, 
Such MCCs identified in (the denser) regions with a lower Toomre $Q_{\rm eff}$ parameter are unstable against collapse.
% Toomre:
Close resemblance of $Q_{\rm eff}$ and $Q_{\rm star}$ maps indicates that contribution from the stellar component plays an important role in governing the stability of the MCCs against axisymmetric perturbations, especially in the central part of \flower. Similarly, stabilizing effect due to the thickness of its disk is also non-negligible. This illustrates the importance of accounting for stellar contribution and disk thickness when examining the stability of molecular gas structures, especially in relatively evolved and enriched systems at high redshift that are preferentially being observed now.
%
Star formation is expected to take place within its star-forming {\it clumps} and {\it cores} on smaller scales, where energy quickly dissipates. This is consistent with the notion that collapsing structures result from gravitational instability occurring within globally stable structures, which are supported by turbulence and rotation on large scale.
%%
%Low-$Q$ sets the regions that fragments 
% and within those regions where alphavir is low enough is where structures to further collapse to form stars.
%
%The MCCs of \flower are more massive and turbulent compared to the Milky Way, with 
% higher gas mass surface densities and velocity dispersions. 
This also implies that \obs on spatially resolution better than $\simeq$40\,pc are needed to examine the truly star-forming structures (cores), and thus, \SF in the first galaxies.
%
%% end
%Determining the multi-phase ISM properties of early galaxies
%is a critical piece to understanding the evolution and
%assembly history of galaxies, since they set the pace
%for chemical reactions and excitation rates for the coolants in the ISM (and subsequent star formation).
%While observations leveraging the combination of spatial-spectral imaging of
%multi-band continuum and spectral line emission are crucial,
%% we show in this work that prototypical galaxies at the EoR with properties similar to \flower
%% are expected to have molecular gas structures on sub-kpc scales. 
% Higher resolution imaging of molecular gas 
Such resolution is within reach with ALMA, for which 0.01" resolution images are available for the high freq bands. The Next Generation VLA (ngVLA) in the 2030s will enable deeper imaging of the cold gas down to $\sim$\,200\,pc via \aco line emission. % 0.005" highest freq; ~0.05 for 16 GHz; CO10 for z~6
On the other hand, cosmological zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy statistics and are subject to the sub-grid models adopted, serve as a useful tool for examining and making predictions on the morphology and dynamics of the molecular ISM of the first galaxies.
%These predictions can in turn be
%tested with high resolution imaging of the gas content in the first galaxies with telescopes such as ALMA and the ngVLA, thereby
%enabling us to test our findings and refine future simulations to shed light on the physical processes behind
%\SF since the cosmic dark ages.
